\chapter{System Requirements Document}

\section*{Purpose}
This document defines the baseline requirements for the `pq-signature-appliance` proof of concept. The appliance performs ML-DSA-87 signatures on a 64-byte digest using a hardware accelerator in PL controlled by Linux software on the Zynq UltraScale+ PS.

\section{Scope}
The initial system target is a correctness-focused signing appliance with a gRPC network API and a hardware-assisted signing path. The current repository does not yet contain the actual cryptographic core implementation.

\section{Functional Requirements}
\begin{longtable}{p{0.15\textwidth} p{0.75\textwidth}}
\toprule
ID & Requirement \\
\midrule
FR-001 & The system shall expose a gRPC service for pre-hash signing requests over Ethernet. \\
FR-002 & The signing request shall contain exactly one 64-byte digest for the initial PoC API. \\
FR-003 & Linux software on the PS shall validate request shape before accessing hardware. \\
FR-004 & The PS shall communicate with the PL through an AXI-Lite wrapper interface for the PoC. \\
FR-005 & The PL-side signing path shall report busy, done, and error status to software. \\
FR-006 & The system shall return the generated signature and its reported length to the client. \\
FR-007 & The system shall support repeated signing requests during continuous operation testing. \\
FR-008 & The repository shall preserve a clean separation between software, hardware, verification, and planning artifacts. \\
FR-009 & The PoC may use hardcoded key material in PL strictly for controlled bring-up and verification. \\
FR-010 & The design shall allow later replacement of placeholder components with a real ML-DSA core without redefining the external service contract. \\
\bottomrule
\end{longtable}

\section{Non-Functional Requirements}
\begin{longtable}{p{0.15\textwidth} p{0.75\textwidth}}
\toprule
ID & Requirement \\
\midrule
NFR-001 & The initial PoC shall prioritize correctness and observability over maximum throughput. \\
NFR-002 & Architecture and interface changes shall be documented in LaTeX and planning artifacts. \\
NFR-003 & The system shall be structured for incremental multi-agent development. \\
NFR-004 & Hardware and software subsystems shall be independently testable before full integration. \\
NFR-005 & Generated EDA artifacts shall not be treated as source unless explicitly approved. \\
\bottomrule
\end{longtable}

\section{Assumptions and Exclusions}
\begin{itemize}[leftmargin=*]
  \item No third-party ML-DSA implementation is integrated at bootstrap time.
  \item Final key storage, provisioning, and production hardening are out of scope for the initial PoC.
  \item Throughput targets beyond basic modeling are deferred to later project phases.
\end{itemize}
