\chapter{Interface Control Document}

\section*{Purpose}
This chapter defines the baseline interface contracts between the external client and the appliance, and between PS software and PL logic.

\section{External Service Interface}
The appliance exposes a gRPC service named `SigningService` with a `SignPrehash` method. The request contains a single digest field that shall be exactly 64 bytes. The response returns signature bytes, a reported signature length, and a coarse status string for the PoC.

\section{External Interface Semantics}
\begin{itemize}[leftmargin=*]
  \item Requests that do not contain exactly 64 digest bytes shall be rejected in software.
  \item The PoC service processes one signing operation at a time unless future software updates explicitly add queueing.
  \item The service shall surface hardware timeout or error conditions to the client.
\end{itemize}

\section{PS-Software Interface to the Wrapper}
The software stack now uses a selectable backend beneath the common device abstraction:
\begin{itemize}[leftmargin=*]
  \item `fake` backend for local development and deterministic tests
  \item `real` backend for Linux-based user-space MMIO access on the target board
\end{itemize}

Both backends present the same logical device operations: digest write, start, status read, completion wait, signature readback, and clear-status.

\section{PS-PL Interface}
The PS-PL boundary is an AXI-Lite register interface. Software writes command and digest registers, triggers execution, polls or waits for completion, and reads status plus signature output data. The wrapper base address is platform-dependent and supplied through board bring-up configuration rather than hardcoded software constants.

\section{Interface Stability Rules}
\begin{itemize}[leftmargin=*]
  \item Changes to the gRPC message contract require matching updates to proto files and software architecture documentation.
  \item Changes to register-level behavior require matching updates to the wrapper specification and register map.
  \item Future board bring-up shall begin with `STUB` mode before attempting `MLDSA\_OSH` mode.
  \item Future engine integration must preserve the PS-visible control semantics unless explicitly reviewed.
\end{itemize}