\chapter{Detailed Register Map}

\section*{Purpose}
This chapter defines the stabilized AXI-Lite register map implemented by the software fake backend and the RTL wrapper.

\section{Register Summary}
\begin{longtable}{p{0.14\textwidth} p{0.19\textwidth} p{0.12\textwidth} p{0.45\textwidth}}
\toprule
Offset & Name & Access & Description \\
\midrule
0x000 & CONTROL & RW & Start and clear-status control bits. \\
0x004 & STATUS & RO & Idle, busy, done, and error status indicators. \\
0x008 & ERROR\_CODE & RO & Coarse error code for wrapper or engine-detected failure. \\
0x00C & SIG\_LENGTH & RO & Reported signature length in bytes. \\
0x010-0x04C & DIGEST[0:15] & RW & Sixteen 32-bit words holding the 64-byte input digest. \\
0x100-0x1310 & SIG\_DATA[0:1156] & RO & Signature read window sized for the largest supported mode result. Deterministic modes use the leading 128 bytes; ML-DSA-87 uses up to 4627 bytes. \\
\bottomrule
\end{longtable}

\section{Control Register Bits}
\begin{itemize}[leftmargin=*]
  \item Bit 0: `start`
  \item Bit 1: `clear\_status`
\end{itemize}

\section{Status Register Bits}
\begin{itemize}[leftmargin=*]
  \item Bit 0: `idle`
  \item Bit 1: `busy`
  \item Bit 2: `done`
  \item Bit 3: `error`
\end{itemize}

\section{Current Wrapper Semantics}
\begin{itemize}[leftmargin=*]
  \item Software shall fully populate the digest registers before asserting `start`.
  \item `start` shall transition the wrapper to busy when idle.
  \item The adapter completes after a mode-specific delay or reports an engine error.
  \item `clear\_status` clears done and error latches and clears the signature window when the wrapper is idle.
  \item `SIG\_LENGTH` reports the valid byte count for the selected engine mode.
\end{itemize}

\section{Current Engine-Mode Semantics}
\begin{itemize}[leftmargin=*]
  \item `STUB`: `STUBSIG || digest || zero padding`, fixed length 128 bytes, deterministic two-cycle completion.
  \item `CORE\_PLACEHOLDER`: `COREPH || digest || zero padding`, fixed length 128 bytes, deterministic four-cycle completion.
  \item `MLDSA\_OSH`: real imported-core attachment path. The intended success case reports the real ML-DSA-87 signature length of 4627 bytes. In the current local Verilog-only regression flow, this mode is exercised via a documented fallback that reports engine error and zero signature length instead of claiming success.
\end{itemize}

\section{Current Error Codes}
\begin{itemize}[leftmargin=*]
  \item 0: no error
  \item 1: start requested while busy
  \item 2: unsupported register offset touched by the current wrapper implementation
  \item 3: engine adapter error, including the current local `MLDSA\_OSH` fallback path when the real core is not elaborated
\end{itemize}

\section{Compatibility Note}
The register base offsets are unchanged from earlier phases. The only compatible extension in the current phase is the larger readable extent of the `SIG\_DATA` window so the wrapper can expose the full ML-DSA-87 result without changing the logical register-bank structure.