\chapter{Verification and Validation Plan}

\section*{Purpose}
This chapter defines the current verification strategy for the appliance.

\section{Current Verification Goal}
The current phase validates that the external request path, MMIO sequencing, wrapper behavior, engine-adapter behavior, and imported-core integration seam are stable enough to support later target-hardware bring-up.

\section{Verification Layers}
\begin{longtable}{p{0.22\textwidth} p{0.68\textwidth}}
\toprule
Layer & Objective \\
\midrule
Documentation review & Confirm consistency across requirements, architecture, interfaces, planning, prompts, and imported-core inspection notes. \\
Software unit tests & Validate request checking, fake backend state transitions, and local client or service contracts. \\
Wrapper and adapter testbench & Validate reset, digest writes, start-to-busy transition, deterministic completion, signature readback, and all currently supported wrapper-visible engine modes. \\
Imported-core inspection & Confirm that project-owned shim logic is based on the actual imported ML-DSA-OSH sign path rather than assumptions. \\
Mixed-language real-core verification & Add full local simulation or synthesis-backed validation for the imported sign path when tooling permits. \\
Bring-up planning & Prepare PS-to-PL and continuous-operation testing on the real platform without changing the stable wrapper contract. \\
\bottomrule
\end{longtable}

\section{Acceptance Focus for the Current Phase}
The current phase is accepted primarily on interface stability, provenance clarity, inspected-core attachment correctness at the architectural level, and preservation of the existing software contract. It is explicitly not accepted on the basis of already-proven end-to-end cryptographic correctness in the current local tool flow.

\section{Current Test Assets}
\begin{itemize}[leftmargin=*]
  \item Python unit tests for digest validation and fake backend state transitions
  \item Python contract tests for the local client and service path
  \item A SystemVerilog wrapper testbench that covers `STUB`, `CORE\_PLACEHOLDER`, and the documented `MLDSA\_OSH` fallback behavior
  \item Imported-core inspection notes tied to the actual upstream source files
\end{itemize}

\section{Current Limits}
The imported upstream sign path mixes Verilog and VHDL. The current portable local environment does not provide mixed-language elaboration, so full real-core local simulation is still pending. This limitation shall be tracked explicitly rather than hidden behind placeholder success cases.

\section{Key Exit Criteria}
\begin{itemize}[leftmargin=*]
  \item Documented registers and control bits are implemented consistently in software and the wrapper.
  \item Deterministic `STUB` mode remains simulation-backed for regression.
  \item The `MLDSA\_OSH` adapter and shim path is documented against the real imported sign interface.
  \item The software execution path runs without real hardware.
  \item The repository is ready for a follow-on phase that adds real PS or PL bring-up and stronger verification of the imported sign path.
\end{itemize}